\documentclass{minimal}
\begin{document}


Reference List\\

Acemoglu, Daron, Tristan Reed, and James A. Robinson. "Chiefs: Economic development and elite control of civil society in Sierra Leone." Journal of Political Economy 122.2 (2014): 319-368.\\

Baldwin, Kate. The paradox of traditional chiefs in democratic Africa. Cambridge University Press, 2015.\\

Blattman, Christopher, Mathilde Emeriau, and Nathan Fiala. Do anti-poverty programs sway voters? Experimental evidence from Uganda. No. w23035. National Bureau of Economic Research, 2017.\\

De Kadt, Daniel, and Evan S. Lieberman. "Nuanced accountability: Voter responses to service delivery in southern africa." British Journal of Political Science (2017): 1-31.\\

De La, O., and L. Ana. "Do conditional cash transfers affect electoral behavior? Evidence from a randomized experiment in Mexico." American Journal of Political Science 57.1 (2013): 1-14.\\

Golden, Miriam, and Brian Min. "Distributive politics around the world." Annual Review of Political Science 16 (2013): 73-99.\\

Guiteras, Raymond P., and Ahmed Mushfiq Mobarak. "Does Development Aid Undermine Political Accountability? Leader and Constituent Responses to a Large-Scale Intervention." (2016).\\

Hartmann, Felix. "Public Transfers and Incumbent Voting: Experimental Evidence from the Philippines." Work (2017).\\

Healy, Andrew, and Neil Malhotra. "Retrospective voting reconsidered." Annual Review of Political Science 16 (2013): 285-306.\\

Jablonski, Ryan S. "How aid targets votes: the impact of electoral incentives on foreign aid distribution." World Politics66.2 (2014): 293-330.\\

Koter, Dominika. "King makers: Local leaders and ethnic politics in Africa." World Politics 65.2 (2013): 187-232.\\

Moss, Todd J., Gunilla Pettersson Gelander, and Nicolas Van de Walle. "An aid-institutions paradox? A review essay on aid dependency and state building in sub-Saharan Africa." (2006).\\

Sawyer, Edward. "Remove or reform? A case for (restructuring) chiefdom governance in post-conflict Sierra Leone." African Affairs 107.428 (2008): 387-403.\\

Stokes et al. Brokers, Voters, and Clientelism: The Puzzle of Distributive Politics. Cambridge University Press, 2013.\\

\end{document}